`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks

### 2019-07-30 analysis/paper/sim_survey_paper.Rmd
Address Joanne's suggestions

---
title: "`SimSurvey`: an `R` package for simulation testing the design and analysis of fisheries-independent trawl surveys"
subtitle: "ALTERNATE TITLE - `SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, Jonathan Babyn, Fran Mowbray, etc.?"
date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
bibliography: references.bib
csl: ices-journal-of-marine-science.csl
---

@@

Fisheries-independent trawl surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998]. While costly to obtain, this information forms the basis of many stock assessments throughout the world. Efficient and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection. While simulations provide a platform for exploring solutions to this optimization problem, building the necessary simulation framework is not a trivial task given the multi-stage nature of the typical sampling program and the complexity of the population processes we aim to represent. Notably, many fish stocks show size-specific patterns of aggregation, making it difficult to collect biological sub-samples (e.g. lengths and ages) that are representative of the entire target population. This undermines the assumptions of many analyses as sub-samples from such populations tend to exhibit strong positive intracluster correlation [@Aanes2015]. These complexities may explain why simulations that test the full sampling and analytical pathway of fisheries-independent surveys are rare [but see @puerta2019; @schnute2003].

@@

The package has two main components, the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated fish stock distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.

@@

The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtural fish populations with correlation across space, time and age dimensions, 2) simulating surveys with a range of sampling stratigies and intensities, and 3) running stratified analyses of simulated survey data (Table \@ref(tab:fun)). The equations behind these functions are detailed in the sections below. **`SimSurvey`** relies heavily on functions from the **`data.table`** [@dowle2017], **`raster`** [@hijmans2016] and **`plotly`** [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively. Package documentation has been published online using **`pkgdown`** (https://paulregular.github.io/SimSurvey/) and all the source `R` code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey). **`SimSurvey`** can be installed via GitHub using the **`remotes`** package:

@@

$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
Here, numbers at age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e. recruits) vary around a baseline value, $log(N_{1,y}) = log(\mu_r) + \epsilon_{y}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(\mu_Z) + \delta_{a,y}$. The error around the recruitment process was set to follow a random walk, $\epsilon_{y} \sim N(\epsilon_{y - 1}, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$. The covariance across ages and years is controlled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, age}$ and $\varphi_{\delta, year}$, respectively). This structure allows for autocorrelation in process errors across ages and years (i.e. total mortality can be made to be more similar for fish that are closer together in age and time). Abundance at age is then converted to abundance at length using the original von Bertalanffy growth curve [@Cailliet2006]:

@@

Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$. Abundance in discrete length groups, $l_{group}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance at age to abundance at length. Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.

@@

| **Function call**                           | **Description**                                            | **Symbol**                 |
|:------------------------------------------- |:---------------------------------------------------------- | :------------------------- |
| `sim_abundance(`                            | &nbsp;                                                     | &nbsp;                     |
| `+  ages = 1:20,`                           | Ages                                                       | $a$                        |
| `+  years = 1:20,`                          | Years                                                      | $y$                        |
| `+  R = sim_R(mean = 30000000,`             | Mean recruitment                                           | $\mu_r$                    |
| `+            log_sd = 0.5),`               | Standard deviation of log-recruitment                      | $\sigma_r$                 |
| `+  Z = sim_Z(mean = 0.5,`                  | Mean total mortality                                       | $\mu_Z$                    |
| `+            log_sd = 0.2,`                | Standard deviation of total mortality (log)                | $\sigma_Z$                 |
| `+            phi_age = 0.9,`               | Correlation across ages in error around total mortality    | $\varphi_{\delta, age}$    |
| `+            phi_year = 0.5),`             | Correlation across years in error around total mortality   | $\varphi_{\delta, year}$   |
| `+  growth = sim_vonB(Linf = 120,`          | Mean asymptotic length (cm)                                | $l_{\infty}$               |
| `+                    L0 = 5,`              | Length at birth (cm)                                       | $l_0$                      |
| `+                    K = 0.1,`             | Growth rate                                                | $K$                        |
| `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_l$                 |
| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{group}^{N}$            |

@@

Abundance at age and length is simulated using the `sim_abundance` function and a default function call is described in Table \@ref(tab:abundance) along with associated symbols from the equations outlined above. This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`). The package includes the functions `sim_R`, `sim_Z` and `sim_vonB` for simulating recruitment, total mortality and growth, respectively. These functions are "closures" [i.e. functions that contain data and return functions; @wickham2014] and this structure was chosen to avoid the repeated specifications of ages and years. The use of closures also allows users to use their own closures with a similar structure but different underlying formula. Overall, the function provides a simple tool for simulating a range dynamic age-structured populations. For instance, below we simulate a relatively long (object `a`) and short (object `b`) lived species, with default variance, starting abundance and growth settings.

```{r, echo = TRUE, eval = replot}
set.seed(438)
a <- sim_abundance(ages = 1:20,
                   R = sim_R(mean = 3e+07),
                   Z = sim_Z(mean = 0.5))
b <- sim_abundance(ages = 1:6,
                   R = sim_R(mean = 1e+10),
                   Z = sim_Z(mean = 0.8))
```

@@

The package also includes several plotting function for making quick plotly-based [@sievert2018] interactive visuals of the simulated population. For instance, the `plot_surface` function can be used to make quick visuals of matrices contained within the list returned by `sim_abundance`. As an example, we display the abundance at age matrix (object named `N` in the list produced by `sim_abundance`; Figure \@ref(fig:surface)); other names can be supplied to the `mat` argument to visualize a different matrix from the `sim_abundance` list, such as `Z`.

```{r, echo = TRUE, eval = FALSE}
plot_surface(a, mat = "N")
plot_surface(b, mat = "N")
```

@@

pa <- plot_surface(a, mat = "N", scene = "scene1", showscale = FALSE) %>% 
  add_labels(x = 0.05, y = 0.9, text = "a)")
pb <- plot_surface(b, mat = "N", scene = "scene2", showscale = FALSE) %>% 
  add_labels(x = 0.05, y = 0.9, text = "b)")

@@

The next step in the simulation is to distribute the abundance at age matrix simulated using the cohort model throughout a spatial field. Here, a grid of $s$ locations is generated with a resolution of $R$ and a depth gradient. The depth gradient is set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$. The grid is divided into $N_{div}$ divisions and $N_{strat}$ depth-based strata. The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space

@@

Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean depth. This helps simulate depth preferences of particular fish. Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and the two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years. Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, age}$ and $\varphi_{\xi, year}$, respectively. The overall variance of the process is controlled by $\sigma_{\xi}^2$. In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering. Note that $\eta_{a,y,s}$ is forced to sum to one in each year to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model. Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid. A more detailed description of the space-age-year covariance is included in Appendix B.

@@

| **Function call**                                       | **Description**                                       | **Symbol**            |
|:------------------------------------------------------- |:----------------------------------------------------- | :-------------------- |
| `sim_distribution(`                                     | &nbsp;                                                | &nbsp;                |
| `+  sim,`                                               | Simulated population from `sim_abundance`             | &nbsp;                |
| `+  grid = make_grid(x_range = c(-140, 140),`           | Range of grid in the x dimension                      | &nbsp;                |
| `+                   y_range = c(-140, 140),`           | Range of grid in the y dimension                      | &nbsp;                |
| `+                   res = c(3.5, 3.5),`                | Grid resolution (km)                                  | $R$                   |
| `+                   shelf_depth = 200,`                | Shelf depth (km)                                      | $d_{shelf}$           |
| `+                   shelf_width = 100,`                | Shelf width (km)                                      | $w_{shelf}$           |
| `+                   depth_range = c(0, 1000),`         | Depth range (m)                                       | $[d_{min}, d_{max}]$  |
| `+                   n_div = 1,`                        | Number of divisions                                   | $N_{div}$             |
| `+                   strat_breaks = seq(0, 1000, 40)),` | Series of depth breaks for defining strata            | &nbsp;                |
| `+  ays_covar = sim_ays_covar(sd = 2.8,`                | Standard deviation of age-year-space distribution     | $\sigma_{\xi}$        |
| `+                            range = 300,`             | Range of spatial correlation (km)                     | $r$                   |
| `+                            lambda = 1,`              | Smoothness of spatial correlation                     | $\lambda$             |
| `+                            phi_age = 0.5,`           | Correlation across ages in spatial distribution       | $\varphi_{\xi, age}$  |
| `+                            phi_year = 0.9,`          | Correlation across years in spatial distribution      | $\varphi_{\xi, year}$ |
| `+                            group_ages = 5:20),`      | Make space-age-year variance equal across these ages  | &nbsp;                |
| `+  depth_par = sim_parabola(mu = 200,`                 | Depth at which abundance is typically highest (m)     | $\mu_d$               |
| `+                           sigma = 70))`              | Dispersion around depth of peak abundance (m)         | $\sigma_d$            |

@@

The above equations are used in the `make_grid`, `sim_ays_covar` and `sim_parabola` functions, and these functions are used within `sim_distribution` to distribute a population simulated using `sim_abundance` throughout a grid (Table \@ref(tab:distribution)). The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`. If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure is used. The package includes a manually constructed survey grid of NAFO division 3Ps (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid. However, for simplicity, we use `make_grid` to construct a square grid for a default run of `sim_distribution`. Below we generate and plot (Figure \@ref(fig:grid)) a default grid, another grid with the number of strata increased by increasing the number of `strat_splits`, and another with four divisions and a more linear gradient in depth.

@@

In addition to supplying objects produced by `sim_abundance` and `make_grid`, the `sim_distribution` function requires two closures that describe the age-year-space covariance and the relationship with depth. Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by tweaking a few parameters in these closures. Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e. wider range) and exhibits stronger correlation across ages than years. Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years` arguments, respectively, in the `sim_ays_covar` closure. Variance in the size of the clusters can also be modified by changing the `sd` argument in the `sim_ays_covar` function. In other words, these parameters can be modified to control the degree of age-specific clustering and inter-annual site-fidelity exhibited by the simulated population. Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete. Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe [`%>%`; @bache2014] friendly.

@@

The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated. The area of each strata is calculated and this is used to define the number of sets to allocate to each strata under a particular set density, $D_{sets}$. The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:

$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{trawl}}{A_{cell}} q_a )$$
Where $n_{a,y,s}$ is the number of fish sampled by a set, $A_{trawl}$ indicates the area covered by the trawl, $A_{cell}$ is the area of a grid cell, and $q_a$ is the catchability coefficient of each age. Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$. The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the [Simulate Abundance] section. Depending on the number of fish caught, sub-sampling is then conducted. Specifically, a maximum number of lengths are measured per set, $M_{lengths}$, and a maximum number of ages, $M_{ages}$, are sampled per length group, $l_{group}$, per division. Such sub-sampling is common in fisheries-independent surveys as it is costly to sample every fish captured. Age determination is especially time-consuming, which is why age sampling tends to be length-stratified.

@@

| **Function call**                   | **Description**                                                         | **Symbol**     |
|:----------------------------------- |:----------------------------------------------------------------------- | :------------- |
| `sim_survey(`                       | &nbsp;                                                                  | &nbsp;         |
| `+  sim,`                           | Simulated spatial population from `sim_distribution`                    | &nbsp;         |
| `+  n_sims = 1`                     | Number of times to repeat the survey                                    | &nbsp;         |
| `+  q = sim_logistic(k = 2,`        | Steepness of logistic curve of catchability                             | $k$            |
| `+                   x0 = 3),`      | Midpoint of logistic curve of catchability                              | $x_0$          |
| `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (length, width; km) - i.e. area covered by a trawl     | $A_{trawl}$    |
| `+  min_sets = 2`                   | Minimum number of sets to conduct per strata                            | &nbsp;         |
| `+  set_den = 2/1000,`              | Set density (km^-2^)                                                    | $D_{sets}$     |
| `+  lengths_cap = 500,`             | Maximum number of lengths to collect / set                              | $M_{lengths}$  |
| `+  length_group = 1,`              | Length group bin size for age sampling (cm)                             | $l_{group}$    |
| `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / division              | $M_{ages}$     |
| `+  age_sammpling = "stratified")`  | Controls whether age sampling is length `"stratified"` or `"random"`    | &nbsp;         |

@@

The function `sim_survey` can be used to simulate data from one survey over a population created using `sim_distribution`. A default function call is described in Table \@ref(tab:survey)). The `sim_survey` function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and definitions of the design of the survey. Specifically, the `q` argument requires a closure, such as `sim_logistic`, for defining the probability of catching specific age groups, trawl dimensions are defined in the `trawl_dim` argument, and set, length and age sampling effort are defined using the `set_den`, `lengths_cap` and `ages_cap` arguments, respectively. Multiple simulations of the same survey can be run using the `n_sims` argument, however, requesting large numbers of simulations can be computational demanding depending on the processing capacity available. Below we use `sim_survey` to simulate two surveys over a default population, of which one is set-up to have higher sampling effort than the other.

@@

>>>>

---
title: "`SimSurvey`: an `R` package to optimize the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"
author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, Jonathan Babyn, Fran Mowbray, etc.?"
date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
bibliography: references.bib
csl: ices-journal-of-marine-science.csl
---

@@

Fisheries-independent surveys have become a mainstay in the management of dynamic fish stocks as they provide indices of population abundance as well as estimates of various population characteristics such as length and age frequencies [@Pennington1998]. While costly to obtain, this information forms the basis of many stock assessments throughout the world. Efficient and informed management decisions therefore require surveys and analyses that maximize information while minimizing the expense of data collection. While simulations provide a platform for exploring solutions to this optimization problem, building the necessary simulation framework is not a trivial task given the multi-stage nature of the typical sampling program and the complexity of the population processes we aim to represent. Notably, many fish stocks show size-specific patterns of aggregation, making it difficult to collect biological sub-samples (e.g. lengths and ages) that are representative of the entire target population. This undermines the assumptions of many analyses as sub-samples from such populations tend to exhibit strong positive intracluster correlation [@Aanes2015]. These complexities may explain why simulations that test the full sampling and analytical pathway of fisheries-independent surveys are rare [but see @puerta2019; @schnute2003].

@@

The package has two main components, the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated population distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.

@@

The **`SimSurvey`** package was written in the programming language `R` [@R] and it holds a series of functions for 1) simulating the abundance and distribution of virtural fish populations with correlation across space, time and age, 2) simulating surveys with a range of sampling stratigies and intensities, and 3) running stratified analyses of simulated survey data (Table \@ref(tab:fun)). The equations behind these functions are detailed in the sections below. **`SimSurvey`** relies heavily on functions from the **`data.table`** [@dowle2017], **`raster`** [@hijmans2016] and **`plotly`** [@sievert2018] packages for their efficient data processing, geographic and plotting facilities, respectively. Package documentation has been published online using **`pkgdown`** (https://paulregular.github.io/SimSurvey/) and all the source `R` code behind **`SimSurvey`** is available on GitHub (https://github.com/PaulRegular/SimSurvey). **`SimSurvey`** can be installed via GitHub using the **`remotes`** package:

@@

$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
Here, numbers at age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e. recruits) vary around a baseline value, $log(N_{1,y}) = log(\mu_r) + \epsilon_{y}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(\mu_Z) + \delta_{a,y}$. The error around the recruitment process was set to follow a random walk, $\epsilon_{y} \sim N(\epsilon_{y - 1}, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$. The covariance across ages and years is controlled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, \mathrm{age}}$ and $\varphi_{\delta, \mathrm{year}}$, respectively). This structure allows for autocorrelation in process errors across ages and years (i.e. total mortality can be made to be more similar for fish that are closer together in age and time). Abundance at age is then converted to abundance at length using the original von Bertalanffy growth curve [@von1938]:

@@

Where $l_{\infty}$ is the mean asymptotic length, $l_0$ is length at birth, $K$ is the growth rate parameter and the error is assumed to follow the normal distribution, $\varepsilon \sim N(0, \sigma_{l}^2)$. Abundance in discrete length groups, $l_{\mathrm{group}}^{N}$, was determined by using this formula to calculate the probability of being within a specific length group given age and then using the resultant length-age-key to convert abundance at age to abundance at length. Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.

@@

| **Function call**                           | **Description**                                            | **Symbol**                          |
|:------------------------------------------- |:---------------------------------------------------------- | :---------------------------------- |
| `sim_abundance(`                            | &nbsp;                                                     | &nbsp;                              |
| `+  ages = 1:20,`                           | Ages                                                       | $a$                                 |
| `+  years = 1:20,`                          | Years                                                      | $y$                                 |
| `+  R = sim_R(mean = 30000000,`             | Mean recruitment                                           | $\mu_r$                             |
| `+            log_sd = 0.5),`               | Standard deviation of log-recruitment                      | $\sigma_r$                          |
| `+  Z = sim_Z(mean = 0.5,`                  | Mean total mortality                                       | $\mu_Z$                             |
| `+            log_sd = 0.2,`                | Standard deviation of total mortality (log)                | $\sigma_Z$                          |
| `+            phi_age = 0.9,`               | Correlation across ages in error around total mortality    | $\varphi_{\delta, \mathrm{age}}$    |
| `+            phi_year = 0.5),`             | Correlation across years in error around total mortality   | $\varphi_{\delta, \mathrm{year}}$   |
| `+  growth = sim_vonB(Linf = 120,`          | Mean asymptotic length (cm)                                | $l_{\infty}$                        |
| `+                    L0 = 5,`              | Length at birth (cm)                                       | $l_0$                               |
| `+                    K = 0.1,`             | Growth rate                                                | $K$                                 |
| `+                    log_sd = 0.1,`        | Standard deviation of the von Bertalanffy growth curve     | $\sigma_l$                          |
| `+                    length_group = 3))`   | Length group bin size for abundance at length (cm)         | $l_{\mathrm{group}}^{N}$            |

@@

Abundance at age and length is simulated using the `sim_abundance` function and a default function call is described in Table \@ref(tab:abundance) along with associated symbols from the equations outlined above. This function has a simple structure and requires the specification of a series of `ages` and `years` along with functions known as "closures", such as `sim_R`, `sim_Z` and `sim_vonB`, for simulating recruitment (`R`), total mortality (`Z`) and growth (`growth`), respectively. Closures are functions that contain data and return functions [@wickham2014] and this structure was chosen to avoid the repeated specifications of ages and years. The use of closures also allows users to construct and use their own closures with a similar structure but different underlying formula. Overall, the function provides a simple tool for simulating a range dynamic age-structured populations. For instance, below we provide examples where we simulate a relatively long and short lived species (note that default variance, starting abundance and growth settings were used in both simulations).

```{r, echo = TRUE, eval = replot}
set.seed(438)
long <- sim_abundance(ages = 1:20,
                      R = sim_R(mean = 3e+07),
                      Z = sim_Z(mean = 0.2))
short <- sim_abundance(ages = 1:6,
                       R = sim_R(mean = 1e+10),
                       Z = sim_Z(mean = 0.8))
```

@@

The package also includes several plotting functions for making quick plotly-based [@sievert2018] interactive visuals of the simulated population. For instance, the `plot_surface` function can be used to make quick visuals of matrices contained within the list returned by `sim_abundance`. As an example, we display the abundance at age matrix (object named `N` in the list produced by `sim_abundance`; Figure \@ref(fig:surface)); other names can be supplied to the `mat` argument to visualize a different matrix from the `sim_abundance` list, such as `Z`.

```{r, echo = TRUE, eval = FALSE}
plot_surface(long, mat = "N")
plot_surface(short, mat = "N")
```

@@

pa <- plot_surface(long, mat = "N", scene = "scene1", showscale = FALSE) %>% 
  add_labels(x = 0.05, y = 0.9, text = "a)")
pb <- plot_surface(short, mat = "N", scene = "scene2", showscale = FALSE) %>% 
  add_labels(x = 0.05, y = 0.9, text = "b)")

@@

The next step in the simulation is to distribute the abundance at age matrix simulated using the cohort model throughout a spatial field. Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$. The grid is divided into $N_{\mathrm{div}}$ divisions (e.g. NAFO or ICES divisions) and $N_{\mathrm{strat}}$ depth-based strata. The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space

@@

Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or dispersion of abundance around the mean depth. This helps simulate depth preferences of particular fish. Residual noise $\xi_{a,y,s}$ is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation within ages and years, and the two dimension AR1 age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years. Spatial correlations are controlled by a smoothing ($\lambda$) and a scaling parameter ($\kappa$) [here $\kappa$ is approximated from the range parameter ($r$), $\kappa = \sqrt{8 \lambda} / r$; @Blangiardo2015] and correlation across ages and years is controlled by $\varphi_{\xi, \mathrm{age}}$ and $\varphi_{\xi, \mathrm{year}}$, respectively. The overall variance of the process is controlled by $\sigma_{\xi}^2$. In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering. Note that $\eta_{a,y,s}$ is forced to sum to one in each year to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model. Also note that numbers were rounded such that a discrete number of fish are in each cell of the grid. A more detailed description of the space-age-year covariance is included in [*Appendix B*][Appendix B: Age-year-space covariance].

@@

| **Function call**                                       | **Description**                                              | **Symbol**                              |
|:------------------------------------------------------- |:------------------------------------------------------------ | :-------------------------------------- |
| `sim_distribution(`                                     | &nbsp;                                                       | &nbsp;                                  |
| `+  sim,`                                               | Simulated population from `sim_abundance`                    | &nbsp;                                  |
| `+  grid = make_grid(x_range = c(-140, 140),`           | Range of grid in the x dimension                             | &nbsp;                                  |
| `+                   y_range = c(-140, 140),`           | Range of grid in the y dimension                             | &nbsp;                                  |
| `+                   res = c(3.5, 3.5),`                | Grid resolution in x and y dimensions (km) - i.e. cell area  | $A_{\mathrm{cell}}$                     |
| `+                   shelf_depth = 200,`                | Shelf depth (m)                                              | $d_{\mathrm{shelf}}$                    |
| `+                   shelf_width = 100,`                | Shelf width (km)                                             | $w_{\mathrm{shelf}}$                    |
| `+                   depth_range = c(0, 1000),`         | Depth range from coast to slope (m)                          | $[d_{\mathrm{min}}, d_{\mathrm{max}}]$  |
| `+                   n_div = 1,`                        | Number of divisions                                          | $N_{\mathrm{div}}$                      |
| `+                   strat_breaks = seq(0, 1000, 40)),` | Series of depth breaks for defining strata                   | &nbsp;                                  |
| `+  ays_covar = sim_ays_covar(sd = 2.8,`                | Standard deviation of age-year-space distribution            | $\sigma_{\xi}$                          |
| `+                            range = 300,`             | Range of spatial correlation (km)                            | $r$                                     |
| `+                            lambda = 1,`              | Smoothness of spatial correlation                            | $\lambda$                               |
| `+                            phi_age = 0.5,`           | Correlation across ages in spatial distribution              | $\varphi_{\xi, \mathrm{age}}$           |
| `+                            phi_year = 0.9,`          | Correlation across years in spatial distribution             | $\varphi_{\xi, \mathrm{year}}$          |
| `+                            group_ages = 5:20),`      | Make space-age-year variance equal across these ages         | &nbsp;                                  |
| `+  depth_par = sim_parabola(mu = 200,`                 | Depth at which abundance is typically highest (m)            | $\mu_d$                                 |
| `+                           sigma = 70))`              | Dispersion around depth of peak abundance (m)                | $\sigma_d$                              |

@@

The above equations are used in the `make_grid`, `sim_ays_covar` and `sim_parabola` functions, and these functions are used within `sim_distribution` to distribute a population simulated using `sim_abundance` throughout a grid (Table \@ref(tab:distribution)). The output from `make_grid` is a raster object [@hijmans2016] with four layers: `depth`, `cell`, `division` and `strat`. If a more detailed and realistic grid is required, users can manually generate their own survey grid using real data and this grid can be supplied as a raster to `sim_distribution` if the same structure is used. The package includes a manually constructed survey grid of NAFO Subdivision 3Ps (named `survey_grid`) and the data-raw folder in the GitHub directory includes the data and code used to construct this grid. However, for simplicity, we use `make_grid` to construct a square grid for a default run of `sim_distribution`. Below we generate and plot (Figure \@ref(fig:grid)) a default grid, another grid with the number of strata increased by increasing the number of `strat_splits`, and another with four divisions and a linear depth gradient (the sigmoid curve is forced to be linear when `shelf_width` is set to zero).

@@

In addition to supplying objects produced by `sim_abundance` and `make_grid`, the `sim_distribution` function requires two closures that describe the age-year-space covariance and the relationship with depth. Here we use `sim_ays_covar` and `sim_parabola` to control these relationships and a wide range of age and year specific distributions can be obtained by tweaking a few parameters in these closures. Below we run a default `sim_distribution` call, which generates a population that forms tight clusters that are more strongly correlated across years than ages, and another call that generates a population that is more diffuse (i.e. wider `range`) and exhibits stronger correlation across ages than years (i.e. lower `phi_year` and higher `phi_age`). Distributions can also be forced to be the same across ages and years by using the `group_ages` and `group_years` arguments, respectively, in the `sim_ays_covar` closure. Variance in the size of the clusters can also be modified by changing the `sd` argument in the `sim_ays_covar` function. In other words, these parameters can be modified to control the degree of age-specific clustering and inter-annual site-fidelity exhibited by the simulated population. Note that the resolution of the default grid is high and, as such, the simulations below may take minutes to complete. Also note that the key functions in the **`SimSurvey`** package have been set-up to be pipe [`%>%`; @bache2014] friendly.

@@

The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated. The area of each strata is calculated and this is used to define the number of sets to allocate to each strata under a particular set density, $D_{\mathrm{sets}}$. The allocated number of cells are randomly selected in each strata and the number of fish caught in each set is calculated by applying binomial sampling of the fish in each sampled cell by the proportion of the area covered by the trawl and the catchability of each age:

$$n_{a,y,s} \sim Bin(N_{a,y,s}, \frac{A_{\mathrm{trawl}}}{A_{\mathrm{cell}}} q_a )$$

Where $n_{a,y,s}$ is the number of fish of age `a` in year `y` sampled by a set at location `s`, $A_{\mathrm{trawl}}$ indicates the area covered by the trawl, $A_{\mathrm{cell}}$ is the area of a grid cell, and $q_a$ is the catchability coefficient of each age. Here, catchabilities were defined using a logistic curve controlled by a steepness, $k$, and midpoint parameter, $x_0$. The lengths of the fish sampled by the set are then simulated using the von Bertalanffy growth equation found above in the *[Simulate abundance]* section. Depending on the number of fish caught, sub-sampling is then conducted. Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division. Such sub-sampling is common in fisheries-independent surveys as it is costly to sample every fish captured. Age determination is especially time-consuming, which is why otoliths for age-determination tend to be sub-sampled by length-bin to obtain a representative age sample across a wider range of lengths than would be obtained via random sampling.

@@

| **Function call**                   | **Description**                                                         | **Symbol**              |
|:----------------------------------- |:----------------------------------------------------------------------- | :---------------------- |
| `sim_survey(`                       | &nbsp;                                                                  | &nbsp;                  |
| `+  sim,`                           | Simulated spatial population from `sim_distribution`                    | &nbsp;                  |
| `+  n_sims = 1`                     | Number of times to repeat the survey                                    | &nbsp;                  |
| `+  q = sim_logistic(k = 2,`        | Steepness of logistic curve of catchability                             | $k$                     |
| `+                   x0 = 3),`      | Midpoint of logistic curve of catchability                              | $x_0$                   |
| `+  trawl_dim = c(1.5, 0.02),`      | Trawl dimensions (length, width; km) - i.e. area covered by a trawl     | $A_{\mathrm{trawl}}$    |
| `+  min_sets = 2`                   | Minimum number of sets to conduct per strata                            | &nbsp;                  |
| `+  set_den = 2/1000,`              | Set density (km^-2^)                                                    | $D_{\mathrm{sets}}$     |
| `+  lengths_cap = 500,`             | Maximum number of lengths to collect / set                              | $M_{\mathrm{lengths}}$  |
| `+  length_group = 1,`              | Length group bin size for age sampling (cm)                             | $l_{\mathrm{group}}$    |
| `+  ages_cap = 10,`                 | Maximum number of ages to sample / length group / division              | $M_{\mathrm{ages}}$     |
| `+  age_sammpling = "stratified")`  | Controls whether age sampling is length `"stratified"` or `"random"`    | &nbsp;                  |

@@

The function `sim_survey` can be used to simulate data from one survey over a population created using `sim_distribution`. A default function call is described in Table \@ref(tab:survey)). The `sim_survey` function simulates the sampling process of the survey and, as such, requires a closure for defining catchability as a function of age and definitions of the design of the survey. Specifically, the `q` argument requires a closure, such as `sim_logistic`, for defining the probability of catching specific age groups, trawl dimensions are defined in the `trawl_dim` argument, and set, length and age sampling effort are defined using the `set_den`, `lengths_cap` and `ages_cap` arguments, respectively. Multiple simulations of the same survey can be run using the `n_sims` argument, however, requesting large numbers of simulations can be computational demanding depending on the processing capacity available. Below we use `sim_survey` to simulate two surveys over a default population, of which one is set-up to have higher set density (`set_den`) than the other.

@@

# Appendix C: Stratified analysis equations
